\section*{Problem 4}

Decide if the following sets are linearly independent or not. 

\subsection*{Part A.}

\[V = \left\{ 
		\genpmat{1\\-2\\4},
		\genpmat{-2\\4\\-8}
\right\}\]

This is not linearly independent. The second vector can be rewritten as: 

\[\vec{v_2} = -2 \cdot \vec{v_1}\]

\subsection*{Part B.}

\[
	V = 
\left\{ 
	\genpmat{1\\2\\3\\4},
	\genpmat{1\\0\\1\\0},
	\genpmat{2\\2\\4\\0}
\right\}
\]

To show linear independence, or a lack thereof, we will find the number of solutions to: 

\[\alpha \vec{v_1} + \beta \vec{v_2} + \gamma \vec{v_3} = \vec{0}\]

Row reduction here is tempting but entirely unnecessary and unhelpful. 

Turning this directly into a system equations gives us:

\[
	\begin{aligned}
	x_1 + x_2 + 2x_3 &= 0 \\
	2x_1 + 2x_3 &= 0 \\
	3x_1 + x_2 + 4x_3 &= 0 \\
	4x_1 &= 0
	\end{aligned}
\]

From here we directly get $x_1 = 0$. Plugging that into our other equations gives us:

\[
	\begin{aligned}
	x_2 + 2x_3 &= 0 \\
	2x_3 &= 0 \\
	x_2 + 4x_3 &= 0
	\end{aligned}
\]

Here we get $x_3 = 0$, and it follows from the basic substitution afterwards that $x_2 = 0$ as well. 

So, the only unique solution to this problem is:

\[x_1 = x_2 = x_3 = 0\]

Thus, this set of vectors is linearly independent. 

\pagebreak
\subsection*{Part C.}

\[
V = 
\left\{	
	\genpmat{1\\2\\3},
	\genpmat{1\\0\\1},
	\genpmat{2\\2\\4},
	\genpmat{-1\\0\\3}
\right\}
\]

We can do some work on this matrix to make it easier to work with. 

I will convert this to an augmented matrix and do the following matrix operation:

\[
r_3 \to r_3 - 2r_2
\]

This results in:

\[
	\genpmat{
		1&1&2&-1 &\bigm| & 0 \\
		2&0&2&0 &\bigm| & 0\\
		-1&0&0&3 &\bigm| & 0
	}
\]

From here we set $r_1 \to r_1 + r_3$
\[
	\genpmat{
		0&1&2&2 &\bigm| & 0\\
		2&0&2&0 &\bigm| & 0\\
		-1&0&0&3&\bigm| & 0
	}
\]

Converting back to a system gives us:

\[
	\begin{aligned}
		x_2 + 2x_3 + 2x^4 &= 0 \\
		2x_1 + 2x_3 &= 0 \\
		-x_1 + 3x_4 &= 0
	\end{aligned}
\]

Solving this system gives us the following solutions:

\[
	\begin{aligned}
		x_1 &= 3x_4 \\
		x_3 &= -3x_4 \\
		x_2 &= 4x_4
	\end{aligned}
\]

A solution to this is all variables being equal to zero, though that certainly is not the only solution to this system of equations. As such, this is not linearly independent. 
