---
title: "Homework 8"
author: "Brady Lamson"
format: pdf
editor: visual
---

```{r, include=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
```


# Section 12.1

## Problem 7

- $y = $ 28-day standard cured strength (psi)

- $x = $ accelerated strength (psi)

$$
y = 1800 + 1.3x
$$

### Part A

What is the expected value of 28-day strength when accelerated strength is 2500?

```{r}
x <- 2500

1800 + (1.3*x)
```

### Part B

By how much can we expect 28-day strength to change when accelerated strength increases by 1 psi?

The slope, so we can expect it to increase by 1.3 psi.

### Part C

By how much can we expect 28-day strength to change when accelerated strength increases by 100 psi?

+130 psi

### Part D

By how much can we expect 28-day strength to change when accelerated strength increases by 100 psi?

-130

# Section 12.2

## Problem 15

### Part B

Is the value of strength completely and uniquely determined by the value of MOE? Explain.

I would wager no. If that was the case there would be no up and down fluctuation in strength as MOE increases. Obviously other things have an effect as well.

### Part C

$$
y = 3.2925 + 0.10748x
$$
Where y is the strength and x is the modus of elasticity

```{r}
x <- 40
y <- 3.2925 + (0.10748*x)
y
```

I wouldn't feel comfortable predicting a value for strength off of a MOE of 100. As we have no values that high we can't determine this relationship continues linearly.

### Part D

$$
\begin{aligned}
SSE &= 18.736 \\
SST &= 71.605 \\
r^2 &= 0.738
\end{aligned}
$$

They do indicate that the regression does a good job. If more of the total sum of squares was made up by the error and the $r^2$ was lower I would not say this. 

## Problem 16

```{r}
df <- tibble(
  rainfall = c(5,12,14,17,23,30,40,47,55,67,72,81,96,112,127),
  runoff = c(4,10,13,15,15,25,27,46,38,46,53,70,82,99,100)
)

df %>%
  ggplot(aes(x=rainfall, y=runoff)) +
  geom_point() +
  theme_bw()
```

The scatterplot does support the use of a simple linear regression model.

### Part B

```{r}
reg <- lm(runoff ~ rainfall, data = df)
reg %>% summary()
```

Point estimate of the slope is 0.82697. Point estimate of the intercept is -1.12830. 

### Part C

```{r}
predict.lm(reg, newdata = data.frame(rainfall = c(50)))
```

### Part D

```{r}
reg %>% anova()
```

$$
\sigma = \sqrt{\frac{SSE}{n-2}} = \sqrt{\frac{27.5}{13}} \approx 1.45
$$

### Part E

0.9753